---
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
  
---

```{r include=FALSE}
library(tidyverse)
```

# Data Source

### Shooting Incident from 2006 to 2022 dataset

  We choose the NYPD Shooting Incident Dataset as our main dataset in the study. This shooting incident dataset contains every shooting incident that occurred in NYC with years ranging from 2006 to 2022. Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included.

```{r echo=TRUE, message = FALSE}
# Read shooting Incident from 2006 to 2022
NYPD_Shooting_Incident_2006_2022 = 
  read_csv("data/NYPD_Shooting_Incident_Data__Historic__20231120.csv") |>
  janitor::clean_names() 
```

There are a total of 21 variables. Out of which, we consider the variables and description is listed below:

* `INCIDENT_KEY`: Randomly generated persistent ID for each arrest.
* `OCCUR_DATE`: Exact date of the shooting incident.
* `OCCUR_TIME`: Exact time of the shooting incident.
* `BORO`: Borough where the shooting incident occurred.
* `PRECINCT`: Precinct where the shooting incident occurred.
* `JURISDICTION_CODE`: Jurisdiction where the shooting incident occurred. Jurisdiction codes 0(Patrol), 1(Transit) and 2(Housing) represent NYPD whilst codes 3 and more represent non NYPD jurisdictions.
* `PERP_AGE_GROUP`: Perpetrator’s age within a category.
* `PERP_SEX`: Perpetrator’s sex description.
* `PERP_RACE`: Perpetrator’s race description.
* `VIC_AGE_GROUP`: Victim’s age within a category.
* `VIC_SEX`: Victim’s sex description.
* `VIC_RACE`: Victim’s race description.
* `X_COORD_CD`: Midblock X-coordinate for New York State Plane Coordinate System, Long Island Zone, NAD 83, units feet (FIPS 3104).
* `Y_COORD_CD`: Midblock Y-coordinate for New York State Plane Coordinate System, Long Island Zone, NAD 83, units feet (FIPS 3104).
* `Latitude`: Latitude coordinate for Global Coordinate System, WGS 1984, decimal degrees (EPSG 4326).
* `Longitude`: Longitude coordinate for Global Coordinate System, WGS 1984, decimal degrees (EPSG 4326).


### Shooting incident data for 2023

  This dataset lists of every shooting incident that occurred in NYC during the current calendar year. As a supplement to the primary dataset, this dataset added 2023 shooting incidents up to September. Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included.

```{r echo=TRUE, message=FALSE}
# Read shooting incident data for 2023
NYPD_Shooting_Incident_2023 = 
  read_csv("data/NYPD_Shooting_Incident_Data__Year_To_Date__20231129.csv") |>
  janitor::clean_names() |>
  select(-new_georeferenced_column, -statistical_murder_flag)
```

  There are a total of 21 variables. Out of which, the variables and description we considered are same as the given list above.

# Data Processing and Cleaning

### Merge the two datasets
* We used the `bind_rows` to bind these two dataste by row, making a longer result for the whole shooting incidents happened from 2006 to 2023.

### Adjustment of data format
* Separate the year, month and day in `occur_date` into three variables. Also, Separate the hours, minutes, and seconds in "occup_time" into three variables for subsequent statistical analysis.

### Data cleansing
* Remove variables that will not be used in subsequent statistical analysis of the data: `minute`, `second`, `loc_of_occur_desc`, `loc_classfctn_desc`, ` location_desc`.

* The dataset is cleaned and merged in the following code.
```{r}
NYPD_Shooting_Incident_cleaned = 
  bind_rows(NYPD_Shooting_Incident_2006_2022, NYPD_Shooting_Incident_2023) |>
  separate(occur_date, into = c("month", "day", "year"), sep = "/") |>
  separate(occur_time, into = c("hour", "minute", "second"), sep = ":") |>
  select(-minute, -second, -loc_of_occur_desc, -loc_classfctn_desc, -location_desc)
```




