---
title: "Statistical Analyses"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide

---

```{r include=FALSE}
library(tidyverse)
library(readr)
library(ggplot2)
library(ggmap)
library(sf)
library(BSDA)

# Read shooting Incident from 2006 to 2022
NYPD_Shooting_Incident_2006_2022 = 
  read_csv("data/NYPD_Shooting_Incident_Data__Historic__20231120.csv") |>
  janitor::clean_names() 

# Read shooting incident data for 2023
NYPD_Shooting_Incident_2023 = 
  read_csv("data/NYPD_Shooting_Incident_Data__Year_To_Date__20231129.csv") |>
  janitor::clean_names() |>
  select(-new_georeferenced_column, -statistical_murder_flag)

NYPD_Shooting_Incident_cleaned = 
  bind_rows(NYPD_Shooting_Incident_2006_2022, NYPD_Shooting_Incident_2023) |>
  separate(occur_date, into = c("month", "day", "year"), sep = "/") |>
  separate(occur_time, into = c("hour", "minute", "second"), sep = ":") |>
  select(-minute, -second, -loc_of_occur_desc, -loc_classfctn_desc, -location_desc)
```


## Seasonal Comparison of Shooting Incidents

* This statistical analysis aims to investigate whether there are significant differences in the mean number of shooting incidents between seasons in New York City. 
* Visualized using bar graphs. x-axis represents the season (winter, spring, summer, fall), and y-axis represents the number of different shootings in each season. The bar plot provides a visual representation of the distribution of shooting incidents across different seasons. 
```{r}
season = 
  NYPD_Shooting_Incident_cleaned |>
  group_by(month) |>
  distinct(incident_key) |>
  summarise(count = n()) |>
  mutate(season = case_match(
    month,
    "01" ~ "winter",
    "02" ~ "winter",
    "03" ~ "spring",
    "04" ~ "spring",
    "05" ~ "spring",
    "06" ~ "summer",
    "07" ~ "summer",
    "08" ~ "summer",
    "09" ~ "fall",
    "10" ~ "fall",
    "11" ~ "fall",
    "12" ~ "winter",
  )) 

season |>
  ggplot(aes(x = season, y = count)) +
  geom_col()
```

* The largest gaps were found in the summer and winter seasons, which were selected for subsequent analysis.

### Summer-Winter Incident Comparison.

* Null Hypothesis (H0): There is no difference in mean incident numbers between summer and winter. 
* Alternative Hypothesis (H1): The mean incident numbers between summer and winter are different.

A two-sample z-test was conducted to compare the average of summer and winter shooting incidents.
```{r message=FALSE}
summer_winter = 
  NYPD_Shooting_Incident_cleaned |>
  group_by(month, year) |>
  distinct(incident_key) |>
  filter(!year == 2023) |>
  summarise(count = n()) |>
  mutate(season = case_match(
    month,
    "01" ~ "winter",
    "02" ~ "winter",
    "03" ~ "spring",
    "04" ~ "spring",
    "05" ~ "spring",
    "06" ~ "summer",
    "07" ~ "summer",
    "08" ~ "summer",
    "09" ~ "fall",
    "10" ~ "fall",
    "11" ~ "fall",
    "12" ~ "winter",
  )) |>
  select(-year) |>
  ungroup() |>
  filter(season %in% c("winter", "summer"))

summer = 
  summer_winter |>
  filter(season == "summer") |>
  select(count)
winter = 
  summer_winter |>
  filter(season == "winter") |>
  select(count)
summer_winter_after = 
  bind_cols(summer, winter) |>
  rename(summer = count...1,
         winter = count...2)

z_test = z.test(x = summer_winter_after$summer, y = summer_winter_after$winter, sigma.x = sd(summer_winter_after$summer), sigma.y = sd(summer_winter_after$winter))
z_test
```

z-test yielded the following results:

* Z value: 9.4566

* P-value: < 2.2e-16 (very small, indicating strong evidence against the original hypothesis)

* Confidence interval: (48.79, 74.31)

The results of the z-test for both samples indicate that there is a statistically significant difference between the average number of shootings in the summer and winter. p-value is extremely small, providing strong evidence against the original hypothesis. Means that there is a large difference between the average number of shootings in the summer and winter months. 

##  Proportion of Male Shooting Victims in Brooklyn vs. Staten Island.
This statistical analysis aims to compare the proportion of male shooting victims between the boroughs of Manhattan and Queens in New York City. The goal is to investigate whether there is a significant difference in the proportion of male victims between these two boroughs. 

* Null Hypothesis (H0): The proportion of male shooting victims in Brooklyn is equal to the proportion in Staten Island. 
* Alternative Hypothesis (H1): The proportions of male shooting victims in Brooklyn and Staten Island are different. 

Use a two-sample test for equality of proportions with continuity correction to compare the proportions of male shooting victims in Brooklyn and Staten Island. 
```{r message=FALSE}
prop_df = 
  NYPD_Shooting_Incident_cleaned |>
  select(boro, vic_sex) |>
  group_by(boro, vic_sex) |>
  summarize(sum = n())

num_brook = 
  prop_df |>
  filter(boro == "BROOKLYN")

num_SI = 
  prop_df |>
  filter(boro == "STATEN ISLAND")

num_brook = sum(pull(num_brook, sum))

num_SI = sum(pull(num_SI, sum))

num_brook_male =
  prop_df |>
  filter(boro == "BROOKLYN") |>
  filter(vic_sex == "M") |>
  pull(sum)

num_SI_male =
  prop_df |>
  filter(boro == "STATEN ISLAND") |>
  filter(vic_sex == "M") |>
  pull(sum)

prop.test(c(num_brook_male, num_SI_male), n = c(num_brook, num_SI))
```

The test results indicate the following:

* X-squared Value: 3.4568

* Degrees of Freedom (df): 1

* P-Value: 0.06299

* Confidence Interval: (-0.0027, 0.0446)

While the p-value is greater than the conventional significance level of 0.05, suggesting that we do not have strong evidence against the null hypothesis. Indicate that the proportions of male victims in Brooklyn and Staten Island are likely to be similar.











